### 2016-08-17 latex/paper.tex
Fit Tc_nondim figure into textwidth
Wheel deformation and schematics

\documentclass{bmd2016p}

@@

%%%%%%%%%%%%%%%% authors %%%%%%%%%%%%%%%
\begin{center}
\normalsize{\bf{M. Ford$^{*}$, J.M. Papadopoulos$^\#$, 
            O. Balogun$^\dag$}}
\end{center}

@@

\section{DEFORMATION OF THE WHEEL}
...Tensegrity
	\begin{equation}\label{eq:TN}
	N_r = \frac{n_sT}{2\pi}
	\end{equation}

@@

@@

\section{BUCKLING DUE TO INTERNAL FORCES}
The solution to Eqn. \ref{eq:Tc_eq} can be written in non-dimensional form as
	\begin{equation}\label{eq:Tc_nondim}
	\frac{T_{c,n}}{T_e} = \left( \frac{\tilde{u}n^2(n^2-1)^2}{1+\tilde{\mu}n^2} + \lambda \right) \left(\frac{1}{n^2-1} \right)
	\end{equation}
where the ``Euler tension'' $T_e=2\pi EI_{22}/n_sR^2$ is the tension that would produce a compressive force in the rim equal to the classical buckling load for a straight beam-column of length $2\pi R$. The parameter $\tilde{\mu} = \widetilde{GJ}/EI_{22}$ is the ratio of the effective torsional stiffness to the lateral bending stiffness. The parameter $\lambda = \bar{k}_{uu}^{el}R^4/EI_{22}$ is the ratio of the spoke system stiffness $\bar{k}_{uu}^{el}R$ to the rim bending stiffness $EI_{22}/R^3$.

>>>>

\documentclass{bmd2016p}
\usepackage{epstopdf}

@@

%%%%%%%%%%%%%%%% authors %%%%%%%%%%%%%%%
\begin{center}
\normalsize{\bf{Matthew Ford$^{*}$, Jim M. Papadopoulos$^\#$, 
            Oluwaseyi Balogun$^\dag$}}
\end{center}

@@

\section{DEFORMATION OF THE WHEEL}
A typical bicycle wheel (Fig. \ref{fig:schematic} consists of a slender rim connected to a hub by means of a system of very slender metal spokes. The spokes are usually threaded at the rim side into spoke nipples which are inserted through the rim from the outside. The spokes are kept under tension by the rim. Without pretension, the spokes would immediately buckle under a very small load due to their high aspect ratio.

\begin{figure}[!h]
\centering
\includegraphics[scale=1.0]{figures/bmd_figures-06.eps}
\caption{caption}
\label{fig:schematic}
\end{figure}

To hold the spokes under tension, the rim must be under compression. An accurate approximation\footnote{95\% accurate for 6 spokes, $\alpha=10^{\circ}$, $\beta=10^{\circ}$. Increasing the number of spokes only improves the estimate.} was given by Sharp~\cite{Sharp1977a} by considering a force balance on the top (or bottom) half of the rim.
	\begin{equation}\label{eq:TN}
	N_r = \frac{n_sT}{2\pi}
	\end{equation}
The spokes act like guy-wires, both supporting the rim radially and stabilizing the rim laterally. However, if the spoke tension exceeds a critical value $T_c$, the compression in the rim causes the wheel to buckle laterally, bending and twisting out of its initial plane, as shown in Fig. \ref{fig:schematic} (c).

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=1.0]{figures/bmd_figures-07.eps}
\caption{Deformation of a rim cross-section. The shear center $S\rightarrow S'$ translates by the vector $\bm{u} = u\hat{\bm{e}}_1 + v\hat{\bm{e}}_2 + w\hat{\bm{e}}_3$ and rotates through an angle $\phi$.}
\label{fig:def}
\end{figure}

@@

\section{BUCKLING DUE TO SPOKE TENSION}
The solution to Eqn. \ref{eq:Tc_eq} can be written in non-dimensional form as
	\begin{equation}\label{eq:Tc_nondim}
	\frac{T_{c,n}}{T_e} = \left( \frac{\tilde{u}n^2(n^2-1)^2}{1+\tilde{\mu}n^2} + \lambda \right) \left(\frac{1}{n^2-1} \right)
	\end{equation}
where the ``Euler tension'' $T_e=2\pi EI_{22}/n_sR^2$ is the tension that would produce a compressive force in the rim equal to the classical buckling load for a straight beam-column of length $2\pi R$. The parameter $\tilde{\mu} = \widetilde{GJ}/EI_{22}$ is the ratio of the effective torsional stiffness to the lateral bending stiffness. The parameter $\lambda = \bar{k}_{uu}^{el}R^4/EI_{22}$ is the ratio of the spoke system stiffness $\bar{k}_{uu}^{el}R$ to the rim bending stiffness $EI_{22}/R^3$.

\begin{figure}[!ht]
\centering
\includegraphics[scale=1.0]{figures/bmd_figures-05.eps}
\caption{\textbf{(a)} Buckling mode parameter map. Only the first 4 modes are shown. \textbf{(b)}--\textbf{(c)} Normalized buckling tension vs. $\tilde{\mu}$ and $\lambda$. Solid black line = Eqn. \ref{eq:Tc_nondim}, dashed red line = power law, blue stars = finite-element simulations. In (a) $\lambda=10$ while in (b) $\tilde{\mu} = 0.38$.}
\label{fig:Tc_nondim}
\end{figure}

Since $n$ is a discrete variable, there is no closed-form for the minimum $T_{c,n}$ for a given $(\tilde{\mu},\lambda)$. However, two important \textit{approximate} solutions can be derived by replacing the discrete $n$ with a continuous variable $\bar{n}$ and minimizing Eqn. \ref{eq:Tc_nondim} analytically, resulting in two simple power laws for $T_c$.

\subsection{Rims with low torsional stiffness}\label{sec:powerlaw_1}
Many bicycle rims are built from thin-walled open channels which have very low torsional stiffness compared with their bending stiffness. Replacing the discrete mode number $n$ with a continuous variable $\bar{n}$, we adopt the following ansatz: (a) $\bar{n}^2 \gg 1$ and (b) $\tilde{\mu}\bar{n}^2 \ll 1$. Under these conditions, Eqn. \ref{eq:Tc_nondim} becomes
	\begin{equation}\label{eq:Tcn_small_mu}
	T_{c,n} = \frac{2\pi EI_{11}}{n_sR^2} \left( \tilde{\mu}\bar{n}^4 + \frac{\lambda}{\bar{n}^2}\right)
	\end{equation}
Minimizing Eqn. \ref{eq:Tcn_small_mu} with respect to $\bar{n}$ by setting the derivative $dT_{c,n}/d\bar{n}=0$, we obtain the scaling law
	\begin{equation}\label{eq:nbar_small_mu}
	\bar{n} = \left(\frac{\lambda}{2\tilde{\mu}} \right)
	\end{equation}
Substituting Eqn. \ref{eq:nbar_small_mu} into Eqn. \ref{eq:Tc_small_mu} yields, after simplification, a simple power law relationship.
	\begin{equation}\label{eq:Tc_small_mu}
	T_c \approx \frac{11.875}{n_sR^2} \, \left(\bar{k}_{uu}^{el}R^4 \right)^{2/3} \widetilde{GJ}^{1/3}
	\end{equation}
Equation \ref{eq:Tc_small_mu} suggests that when the torsional stiffness is smaller than the bending stiffness, the buckling behavior is dominated by torsion and does not depend explicitly on the bending stiffness at all. Equation \ref{eq:Tc_small_mu} is shown in Fig. \ref{fig:Tc_nondim} next to Equation \ref{eq:Tc_nondim} and finite-element calculations. The power law is strictly less than Equation \ref{eq:Tc_nondim} which conveniently gives a conservative design criterion.

\subsection{Spokes much stiffer than the rim}
A second power law regime appears when $\lambda \gg 1$ and $\tilde{\mu} \sim 1$. Under these conditions, Eqn. \ref{eq:Tc_nondim} becomes
	\begin{equation}\label{eq:Tcn_boef}
	T_{c,n} = \frac{2\pi EI_{11}}{n_sR^2}\left(\bar{n}^2 + \frac{\lambda}{\bar{n}^2} \right)
	\end{equation}
Following the same procedure in section \ref{sec:powerlaw_1} gives a separate power law, after simplification.
	\begin{equation}\label{eq:Tc_boef}
	T_c \approx \frac{4\pi}{n_s} \left(\bar{k}_{uu}^{el}EI_{11} \right)^{1/2} 
	\end{equation}
Substituting Eqn. \ref{eq:TN} for $T_c$ gives $N_{rc}=2\sqrt{\bar{k}_{uu}^{el}EI_{11}}$, which is precisely the buckling load for an infinite, straight beam on an elastic foundation~\cite{Hetenyi1946a}. Figure \ref{fig:Tc_nondim} (c) shows a comparison of Eqn. \ref{eq:Tc_boef} against
